---
title: "Feature Engineering"
author: "Brendan McGuinness"
output: html_document
---

# Data Cleaning

This Notebook will address

-   Additional Data Cleaning
-   Feature Engineering
-   Missingness

Load packages

```{r}
library(dplyr)
library(ggplot2)
library(visdat)
library(tidyr)
```

Load data

**Replace 'file' with your file directory**

```{r}
TE_data <- read.csv("file/TE data.csv")
T_data <- read.csv("file/T data.csv")
RB_data <- read.csv("file/RB data.csv")
G_data <- read.csv("file/G data.csv")
tracking_motion <- read.csv("file/tracking motion.csv")
```

## Data Cleaning

Right/Left side of the ball

```{r}

# First Determine if the TE is facing visitor or home endzone
TE_data <- TE_data %>% 
  mutate(Facing = case_when(
    playDirection == "right" ~ "Visitor",
    playDirection == "left" ~ "Home"
  ))

# If we know what direction they're facing we can determine what side of the qb they are by comparing their position to the ball
TE_data <- TE_data %>%
 mutate(SideBS = case_when(
   Facing == "Visitor" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Right",
     (y_bs - y_fb) > 0 ~ "Left"
     ),
   Facing == "Home" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Left",
     (y_bs - y_fb) > 0 ~ "Right"
     ),))

# Determining if a Tackle is LT or RT
T_data <- T_data %>% 
  mutate(Facing = case_when(
    playDirection == "right" ~ "Visitor",
    playDirection == "left" ~ "Home"
  ))

T_data <- T_data %>%
 mutate(SideBS = case_when(
   Facing == "Visitor" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Right",
     (y_bs - y_fb) > 0 ~ "Left"
     ),
   Facing == "Home" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Left",
     (y_bs - y_fb) > 0 ~ "Right"
     ),))
# Also changing position from logic to 'T'
T_data <- T_data %>%
 mutate(position = case_when(
   position == TRUE ~ "T"
 ))
   
T_data <- T_data %>%
 mutate(TacklePos = case_when(
   SideBS == "Right" ~ "RT",
   SideBS == "Left" ~ "LT"
   ))

# Determining if a Guard is LG or RG

G_data <- G_data %>% 
  mutate(Facing = case_when(
    playDirection == "right" ~ "Visitor",
    playDirection == "left" ~ "Home"
  ))

G_data <- G_data %>%
 mutate(SideBS = case_when(
   Facing == "Visitor" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Right",
     (y_bs - y_fb) > 0 ~ "Left"
     ),
   Facing == "Home" ~ 
     case_when(
     (y_bs - y_fb) < 0 ~ "Left",
     (y_bs - y_fb) > 0 ~ "Right"
     ),))
   
G_data <- G_data %>%
 mutate(TacklePos = case_when(
   SideBS == "Right" ~ "RG",
   SideBS == "Left" ~ "LG"
   ))
```

Tackle Cleaning and Dealing with only 1 tackle

```{r}
T_data <- T_data %>% 
  mutate(super_key = paste(gameId, playId, sep = ""))

G_data <- G_data %>% 
  mutate(super_key = paste(gameId, playId, sep = ""))

T_data <- T_data %>%
  group_by(super_key) %>%
  mutate(NumberofT = n()) %>% 
  ungroup()

G_data <- G_data %>%
  group_by(super_key) %>%
  mutate(NumberofG = n()) %>% 
  ungroup()

# Guard Distance Rank
G_data <- G_data %>% 
  group_by(super_key) %>%
  mutate(GuardDistRank = dense_rank(DistFromBall_Wid)) %>% 
  ungroup()

# Observations where there is only one tackle present
T1_data <- T_data %>% 
  filter(NumberofT == 1)

# Semi joining Guard and one Tackle data set to find out which guards match with the singe Tackles
G_data <- semi_join(G_data, T1_data, by = "super_key")

# Joining T1 and the Guard data inorder to find observations where a guard subsituted for a Tackle that game
T1_G_data <- bind_rows(T1_data, G_data) %>% 
  arrange(super_key)

# Keeping the guard that is the furthest away from the ball (in Tackle Position)
T1_G_data <- T1_G_data %>% 
  group_by(super_key) %>%
  mutate(InTacklePos = case_when(
    NumberofG == GuardDistRank ~ "yes",
    is.na(NumberofG) & is.na(GuardDistRank) ~ "yes",
    TRUE ~ "no"
  )) %>% 
  ungroup()

# Removing all guards that aren't labeled "yes"
T1_G_data <- T1_G_data %>% 
  filter(!(InTacklePos == "no"))

# Check if the guard and the Tackle are on the same side
T1_G_data <- T1_G_data %>% 
  group_by(super_key) %>% 
  mutate(TackleSide = lag(SideBS)) %>%
  mutate(TackleSide = case_when(
    is.na(TackleSide) ~ "None",
    TRUE ~ TackleSide
  )) %>% 
  ungroup()

# Removing if on the same side
T1_G_data <- T1_G_data %>% 
  arrange(super_key) %>% 
  filter(!(TackleSide == SideBS))

# Counting number of Single Tackles
T1_G_data <- T1_G_data %>% 
  group_by(super_key) %>%
  mutate(NumberofT = n()) %>% 
  ungroup()

# Getting rid of the rest of Single Tackles
T1_G_data <- T1_G_data %>% 
  filter(!(NumberofT == 1))

# Changing all Guard Pos to Tackles
T1_G_data <- T1_G_data %>% 
  group_by(super_key) %>% 
  mutate(TacklePos = case_when(
    TacklePos == "RG" ~ "RT",
    TacklePos == "LG" ~ "LT",
    TRUE ~ TacklePos
  )) %>% 
  ungroup()

# Selecting only Necessary Variables
T1_G_data <- T1_G_data %>% 
  select(-c(NumberofG, GuardDistRank, InTacklePos, TackleSide))

# Removing observations with only one tackle present so I can join my new data set
T_data <- T_data %>% 
  filter(!(NumberofT == 1))

# Combining T_data and T1_G_data
T_data <- bind_rows(T_data, T1_G_data)
```

More than 3 Tackles listed on a play

```{r}
# If there are more than 3 tackles on a play, look at the tackles that share a side, the tackle with the largest DistFromBall_Wid is kept. There should only be two tackles for every duplication

# freq represents how many T are on the same side. 

T_data <- T_data %>% 
  group_by(super_key, SideBS) %>% 
  mutate(freq = n()) %>% 
  ungroup()

#If the freq equals that number of T that are on the play, then there are no tackles on the other side of the ball, meaning we must drop these observations
T_data <- T_data %>% 
  filter(!(freq == NumberofT))

# If freq is greater than 1, then there are too many tackles on the play and we need to decide which tackles to keep and which to get rid of
# The tackle with the largest distance from the ball is kept because that is the tackle that would be closest to the TE.

T_data <- T_data %>% 
  group_by(super_key, SideBS) %>%
  mutate(TackleDistRank = dense_rank(DistFromBall_Wid)) %>% 
  mutate(InTacklePos = case_when(
    freq > 1 & TackleDistRank == 2 & freq != 3 ~ "yes",
    freq > 1 & TackleDistRank == 3 ~ "yes",
    TRUE ~ "no"
  )) %>% 
  ungroup()

# dropping observations that aren't set up as tackles (InTacklePos is "no")
T_data <- T_data %>% 
  filter(!(freq > 1 & InTacklePos == "no"))

# Setting up data for merge
T_data <- T_data %>% 
  rename(DistFromBall_Tackle = DistFromBall_Wid) %>% 
  rename(LOS_Tackle = DistFromLOS) %>% 
  select(gameId, playId, SideBS, TacklePos, LOS_Tackle, DistFromBall_Tackle)
```

## Feature Engineering

### TE position

*How do you determine if a TE is in a TE position or not?*

1)  If the TE is within 2 yards of the tackle, then they will be considered a TE

2)  If there are multiple TEs then we will check if they are close to the TE next to them

How to determine position with multiple TEs:

-   `TEposition` classes:

    TEo: If lined up as the outside most TE

    TEi: If one TE is lined up outside of them

    TEii: If two TEs are lined up outside of them

    Not TE: If determined to not be in a TE position

-   Classify each TE based on how far they are from the ball

    TE #1 \~ closest to the ball

    TE #2 \~ next closest

    TE #3 \~ furthest from the ball if there are two other TE

-   Determine position based on logic

    -   TE #3:
        -   Can only be TEo or Not TE
        -   Assuming TE #2 and TE#1 are in position
        -   TEo:
            -   On the same side as either TE and within 2 yards
            -   Not on the same side as either TEs, but within 2 yards of a Tackle
        -   Not TE:
            -   None of the other conditions were met
    -   TE #2:
        -   Can be TEo, TEi, or Not TE
        -   Assuming TE #1 is in position
        -   TEi:
            -   TE #3 exists and is in a TE position, and on the same side
        -   TEo:
            -   TE #3 exists and is in a TE position, TE #2 isn't on the same side of either TE, but is within 2 yards of their Tackle
            -   TE #3 doesn't exists or isn't in a TE position, but TE #2 is on the same side as TE #1, and within 2 yards of TE #1
            -   TE #3 doesn't exists or isn't in a TE position, TE #2 is not on the same side as TE1, but is within 2 yards of their Tackle
        -   Not TE:
            -   None of the other conditions were met
    -   TE #1:
        -   Can be TEo, TEi, TEii, or Not TE
        -   TEii:
            -   Same side of TE #3 and TE #2, and they have been labeled TEo and TEi respectively
        -   TEi:
            -   TE #3 doesn't exists or isn't in a TE position, TE #2 is TEo and TE #1 is on the same side as TE #2
            -   TE #3 and TE #2 are TEos
        -   TEo:
            -   TE #3 and TE #2 are in a TE position, but TE #1 is not on the same side of either, and is within 2 yards of their Tackle
            -   TE #3 doesn't exists or isn't in a TE position, isn't on the same side as TE 2, but is within 2 yards of their Tackle
            -   Neither TE #3 nor TE #2 exist or aren't in a TE position, but is within 2 yards of their Tackle
        -   Not TE:
            -   None of the other conditions were met

```{r}
# Merging Tackle data with TE data
TE_data <- merge(TE_data, T_data, by = c("gameId", "playId", "SideBS"))

TE_data <- TE_data %>% 
  relocate(SideBS, .after = Facing)

# If the TE is closer to the ball width wise than the outside tackle then they are set up as a FB or In motion during the snap. I don't want to be analyzing TEs when they are in this position. SO I'm removing them from the data set.
TE_data <- TE_data %>% 
  filter(!(DistFromBall_Wid < DistFromBall_Tackle))

# Determining if there are multiple TEs on a play
# Creating a super key so it's easier to tell if there are observations that have matching game ids and play ids
TE_data <- TE_data %>% 
  mutate(super_key = paste(gameId, playId, sep = ""))

# Making a new variable where if super_keys match, then there are multiple TE on the play
TE_data <- TE_data %>% 
  mutate(multipleTE = duplicated(super_key) | duplicated(super_key, fromLast = TRUE))

# Creating number of TE on the play
# What rank they are (1 is closest to ball 3 is furthest)
# What the previous rank TE position is
# What the side of the ball the previous rank TEs are
TE_data <- TE_data %>%
  arrange(super_key, nflId) %>% 
  group_by(super_key) %>%
  mutate(
    NumberofTE = n(),
    TErank = rank(DistFromBall_Wid, ties.method = "first"),
    DistTE1 = lag(DistFromBall_Wid, n = 2, order_by = TErank),
    DistTE2 = lag(DistFromBall_Wid, order_by = TErank),
    SideTE1 = lag(SideBS, n = 2, order_by = TErank),
    SideTE2 = lag(SideBS, order_by = TErank)
  ) %>%
  ungroup()

# Determining if 3rd TE is in correct position
TE_data <- TE_data %>% 
  mutate(TEposition = case_when(
    TErank == 3 & 
      (SideBS == SideTE1) &
      (DistFromBall_Wid - DistTE1) < 2 ~ "TEo", 
      # Same side as TE 1, within 2 yards of their Tackle
    TErank == 3 & 
      (SideBS == SideTE2) &
      (DistFromBall_Wid - DistTE2) < 2 ~ "TEo", 
      # Same side as TE 2, within 2 yards of their Tackle
    TErank == 3 &
      (SideBS != SideTE1 & SideBS != SideTE2) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2 ~ "TEo", 
      # not on the same side of either TE, within 2 yards of their Tackle
    TErank == 3 ~ "Not TE"
      # If none of these conditions are met the 3rd TE is not in a TE position
  ))

# Saving if TE3 is a TE or not, so I can determine what position TE2 should be in
TE_data <- TE_data %>% 
  group_by(super_key) %>% 
  mutate(
    PosTE3 = lead(TEposition, order_by = TErank),
    SideTE3 = lead(SideBS, order_by = TErank),
    SideTE1 = lag(SideBS, order_by = TErank),
    DistTE1 = lag(DistFromBall_Wid, order_by = TErank)
    ) %>% 
  ungroup()

# Determining if 2nd TE is in correct position
TE_data <- TE_data %>% 
  mutate(TEposition = case_when(
    TErank == 2 &
      (PosTE3 == "TEo") &
      (SideBS == SideTE3) ~ "TEi", 
      # TE 3 is on the outside, and is same side as TE 3
    TErank == 2 &
      (PosTE3 == "TEo") &
      (SideBS != SideTE1 & SideBS != SideTE3) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2 ~ "TEo", 
      # TE 3 is on the outside, but isn't on the same side as TE 1 or TE3, and within 2 yards of their Tackle
    TErank == 2 &
      (PosTE3 == "Not TE" | is.na(PosTE3)) &
      (SideBS == SideTE1) &
      (DistFromBall_Wid - DistTE1) < 2 ~ "TEo",
      # TE 3 is not a TE or doesn't exist, on the same side as TE1, and within 2 yards of TE 1
    TErank == 2 &
      (PosTE3 == "Not TE" | is.na(PosTE3)) &
      (SideBS != SideTE1) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2 ~ "TEo",
      # TE 3 is not a TE or doesn't exist, not on the same side as TE1, and within 2 yards of their Tackle
    TErank == 2 &
      (SideBS == SideTE1) &
      (DistFromBall_Wid - DistTE1) < 2 ~ "TEo",
      # No TE 3, same side as TE1, and within 2 yards of TE 1
    TErank == 2 ~ "Not TE",
      # If none of these conditions are met the 2nd TE is not in a TE position
    TRUE ~ TEposition
    ))

# Saving if TE2 is a TE or not, so I can determine what position TE1 should be in
TE_data <- TE_data %>% 
  group_by(super_key) %>%
  mutate(
    PosTE2 = lead(TEposition, order_by = TErank),
    SideTE2 = lead(SideBS, order_by = TErank),
    PosTE3 = lead(TEposition, n = 2, order_by = TErank),
    SideTE3 = lead(SideBS, n = 2, order_by = TErank)) %>% 
  ungroup()

# Determining if 1st TE is in correct position
TE_data <- TE_data %>% 
  mutate(TEposition = case_when(
    TErank == 1 &
      (PosTE3 == "TEo" & PosTE2 == "TEi") &
      (SideBS == SideTE3 & SideBS == SideTE3) ~ "TEii",
      # Same side of TE 3 and TE 2, and they have been labeled TEo and TEi
    TErank == 1 &
      (PosTE3 == "Not TE" | is.na(PosTE3)) &
      (PosTE2 == "TEo") &
      (SideBS == SideTE2) ~ "TEi",
      # TE 3 is not a TE, TE 2 is TEo and same side as TE 2
    TErank == 1 &
      (PosTE3 == "TEo" & PosTE2 == "TEo") ~ "TEi",
      # TE 3 and TE 2 are outside TEs
    TErank == 1 &
      (PosTE3 == "TEo" & PosTE2 == "TEi") &
      (SideBS != SideTE3 & SideBS != SideTE3) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2  ~ "TEo",
      # TE 3 and TE 2 are in position, not on the same side as them, and within 2 yards of their Tackle
    TErank == 1 &
      (PosTE3 == "Not TE" | is.na(PosTE3)) &
      (SideBS != SideTE2) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2  ~ "TEo",
      # TE 3 isn't in position or doesn't exist, not on the same side as TE 2,
      # and within 2 yards of their Tackle
    TErank == 1 &
      (PosTE3 == "Not TE" | is.na(PosTE3)) &
      (PosTE2 == "Not TE" | is.na(PosTE2)) &
      (DistFromBall_Wid - DistFromBall_Tackle) < 2  ~ "TEo",
      # If neither TE is in position or exists, and withing 2 yards of their Tackle
    TErank == 1 ~ "Not TE",
      # If none of these conditions are met the 1st TE is not in a TE position
    TRUE ~ TEposition
  ))

TE_data <- TE_data %>% 
  select(-c(blockedPlayerNFLId2, blockedPlayerNFLId3, DistTE1, DistTE2, SideTE1, SideTE2, SideTE3, PosTE3, PosTE2))
```

### TE Alignment

Similar to `receiverAlignment` except in this case we are only showing the alignment for TEs that are in a TE position. If a TE is labeled as "Not TE" then they will not be considered apart of the TE alignment

```{r}
# Determining TE alignment (1x0, 2x1, 0x0, etc)
# For this variable if a TE is labeled as "Not TE" then they won't be considered apart of the TE alignment

# Rank 3 TEs
TE_data <- TE_data %>%
  arrange(super_key, TErank) %>%
  group_by(super_key) %>%
  mutate(
    PosTE1 = case_when(TErank == 3 ~ lag(TEposition, n = 2, order_by = TErank)), 
    SideTE1 = case_when(TErank == 3 ~ lag(SideBS, n = 2, order_by = TErank)),
    PosTE2 = case_when(TErank == 3 ~ lag(TEposition, order_by = TErank)),
    SideTE2 = case_when(TErank == 3 ~ lag(SideBS, order_by = TErank))
  )

# Rank 2 TEs
TE_data <- TE_data %>% 
  mutate(
    PosTE1 = case_when(TErank == 2 ~ lag(TEposition, order_by = TErank), TRUE ~ PosTE1),
    SideTE1 = case_when(TErank == 2 ~ lag(SideBS, order_by = TErank), TRUE ~ SideTE1),
  ) %>% 
  ungroup()

# Making TE Alignment variable
TE_data <- TE_data %>% 
  mutate(TEAlignment = NA)

# Function for if only 1 TE is in position
assign_1TE <- function(data, TEpos, TEside) {
  data %>% 
    mutate(TEAlignment = case_when(
      !!sym(TEpos) == "Not TE" ~ "0x0",
      !!sym(TEside) == "Right" ~ "0x1",
      !!sym(TEside) == "Left" ~ "1x0",
      TRUE ~ TEAlignment
    ))
}

#Function for if 2 TEs are in position
assign_2TE <- function(data, TEside1, TEside2) {
  data %>% 
    mutate(TEAlignment = case_when(
      !!sym(TEside1) == "Left" & !!sym(TEside2) == "Left" ~ "2x0",
      !!sym(TEside1) == "Left" & !!sym(TEside2) == "Right" ~ "1x1",
      !!sym(TEside1) == "Right" & !!sym(TEside2) == "Left" ~ "1x1",
      !!sym(TEside1) == "Right" & !!sym(TEside2) == "Right" ~ "0x2",
      TRUE ~ TEAlignment
    ))
}

#Function for if 3 TEs are in position
assign_3TE <- function(data, TEside1, TEside2) {
  data %>% 
    mutate(TEAlignment = case_when(
      SideBS == "Left" & !!sym(TEside1) == "Left" & !!sym(TEside2) == "Left" ~ "3x0",
      SideBS == "Left" & !!sym(TEside1) == "Left" & !!sym(TEside2) == "Right" ~ "2x1",
      SideBS == "Left" & !!sym(TEside1) == "Right" & !!sym(TEside2) == "Left" ~ "2x1",
      SideBS == "Right" & !!sym(TEside1) == "Left" & !!sym(TEside2) == "Left" ~ "2x1",
      SideBS == "Left" & !!sym(TEside1) == "Right" & !!sym(TEside2) == "Right" ~ "1x2",
      SideBS == "Right" & !!sym(TEside1) == "Right" & !!sym(TEside2) == "Left" ~ "1x2",
      SideBS == "Right" & !!sym(TEside1) == "Left" & !!sym(TEside2) == "Right" ~ "1x2",
      SideBS == "Right" & !!sym(TEside1) == "Right" & !!sym(TEside2) == "Right" ~ "0x3",
      TRUE ~ TEAlignment
    ))
}

# For 3 TE
TE_data <- TE_data %>% 
  mutate(TEAlignment = case_when(
    TErank == 3 & TEposition == "Not TE" & PosTE2 == "Not TE" ~ assign_1TE(TE_data, "PosTE1", "SideTE1")$TEAlignment,
    TErank == 3 & PosTE2 == "Not TE" & PosTE1 == "Not TE" ~ assign_1TE(TE_data, "TEposition", "SideBS")$TEAlignment,
    TErank == 3 & TEposition == "Not TE" ~ assign_2TE(TE_data, "SideTE1", "SideTE2")$TEAlignment,
    TErank == 3 & PosTE2 == "Not TE" ~ assign_2TE(TE_data, "SideTE1", "SideBS")$TEAlignment,
    TErank == 3 & PosTE1 == "Not TE" ~ assign_2TE(TE_data, "SideTE2", "SideBS")$TEAlignment,
    TErank == 3 ~ assign_3TE(TE_data, "SideTE1", "SideTE2")$TEAlignment
  ))

# For 2 TE

#If Alignment has already been assigned by TE3, use that alignment
TE_data <- TE_data %>% 
  group_by(super_key) %>% 
  mutate(TEAlignment = case_when(
    lead(!is.na(TEAlignment), order_by = TErank) ~ lead(TEAlignment, order_by = TErank), #If Alignment has already been assigned by TE3, use that alignment
    TRUE ~ TEAlignment
  )) %>% 
  ungroup()

# Other scenarios
TE_data <- TE_data %>% 
  mutate(TEAlignment = case_when(
    NumberofTE == 2 & TErank == 2 & TEposition == "Not TE" ~ assign_1TE(TE_data, "PosTE1", "SideTE1")$TEAlignment,
    NumberofTE == 2 & TErank == 2 & PosTE1 == "Not TE" ~ assign_1TE(TE_data, "TEposition", "SideBS")$TEAlignment,
    NumberofTE == 2 & TErank == 2 ~ assign_2TE(TE_data, "SideTE1", "SideBS")$TEAlignment,
    TRUE ~ TEAlignment
  ))

# For 1 TE

#If Alignment has already been assigned by TE2, use that alignment
TE_data <- TE_data %>% 
  group_by(super_key) %>% 
  mutate(TEAlignment = case_when(
    lead(!is.na(TEAlignment), order_by = TErank) ~ lead(TEAlignment, order_by = TErank), #If Alignment has already been assigned by TE3, use that alignment
    TRUE ~ TEAlignment
  )) %>% 
  ungroup()

# Only 1 other scenario, if the TE is by itself
TE_data <- TE_data %>% 
  mutate(TEAlignment = case_when(
    NumberofTE == 1 ~ assign_1TE(TE_data, "TEposition", "SideBS")$TEAlignment,
    TRUE ~ TEAlignment
  ))

# Removing unnecessary variables
TE_data <- TE_data %>% 
  select(-c(TErank, SideTE1, SideTE2, PosTE2, PosTE1))
```

### RB Distance from ball

In my exploratory data analysis of RB I noticed that a great indicator for if there was going to be a run or a pass was how far from the ball the RB sets up. This makes intuitive sense since when the RB is about 5 yards away when they're in Shotgun formation, which allows for a higher RB route potential. But if they are about 7 yards away, like in Single back or Pistol formation, there route options are limited and would look more toward running the ball or acting as a decoy. Thus makes a great predictor for determining if a TE will run a route or block since it mostly hinges on if the RB is going to get the ball or not.

*What if there is more than 1 RB?*

We only want the dist of one RB to compare to our TEs, here are the parameters if there is more than one RB:

1)  If the RB has a rush attempt keep
2)  If neither had a rush attempt, keep if RB Distance is over 9 yards away
3)  If both less than 9 yards, keep the RB that's closest
4)  If both more than 9 yards, keep the RB that's furthest

```{r}
# RB data
RB_data_c <- RB_data %>% 
  select(gameId, playId, nflId, hadRushAttempt, DistFromBall_BS)

# Dealing with multiple TEs
RB_data_c <- RB_data_c %>% 
  mutate(super_key = paste(gameId, playId, sep = ""))

# Getting rid of duplicate case
RB_data_c <- RB_data_c %>%
  filter(!(super_key == "20220911034434"))

# Organizing data for cleaning
RB_data_c <- RB_data_c %>%
  arrange(super_key, nflId) %>% 
  group_by(super_key) %>% 
  mutate(NumberofRB = n()) %>%
  mutate(RBrank = rank(nflId)) %>% 
  ungroup()

# If the RB has a rush attempt, keep them over the other RB
RB_data_c <- RB_data_c %>% 
  mutate(Keep = case_when(
    NumberofRB == 2 &
      hadRushAttempt == 1 ~ "TRUE",
    NumberofRB == 2 &
      hadRushAttempt == 0 &
      RBrank == 1 &
      lead(hadRushAttempt == 1) ~ "FALSE",
    NumberofRB == 2 &
      hadRushAttempt == 0 &
      RBrank == 2 &
      lag(hadRushAttempt == 1) ~ "FALSE",
    NumberofRB == 2 ~ "No Rush"
  ))

# If neither had a rush attempt, if over 9 yards away keep
# If both less than 9 yards, keep the RB that's closest
# If both more than 9 yards, keep the RB that's furthest
RB_data_c <- RB_data_c %>% 
  mutate(Keep = case_when(
    Keep == "No Rush" &
      DistFromBall_BS >= 9 &
      RBrank == 1 &
      lead(DistFromBall_BS) < DistFromBall_BS ~ "TRUE",
    Keep == "No Rush" &
      DistFromBall_BS >= 9 &
      RBrank == 2 &
      lag(DistFromBall_BS) < DistFromBall_BS ~ "TRUE",
    Keep == "No Rush" &
      RBrank == 1 &
      lead(DistFromBall_BS) > DistFromBall_BS & 
      lead(DistFromBall_BS < 9) ~ "TRUE",
    Keep == "No Rush" &
      RBrank == 2 &
      lag(DistFromBall_BS) > DistFromBall_BS & 
      lag(DistFromBall_BS < 9) ~ "TRUE",
    Keep == "No Rush" ~ "FALSE",
    TRUE ~ Keep
  ))

RB_data_c <- RB_data_c %>% 
  mutate(Keep = case_when(
    is.na(Keep) ~ "TRUE",
    TRUE ~ Keep))

# Deleting Falses
RB_data_c <- RB_data_c %>% 
  filter(!(Keep == "FALSE"))

# For 3 RBs, only keep if there was a rush attempt
RB_data_c <- RB_data_c %>% 
  filter(!(NumberofRB == 3 & hadRushAttempt != 1))

# Checking to see if there are any more duplicates
RB_data_c <- RB_data_c %>%
  group_by(super_key) %>% 
  mutate(NumberofRB = n()) %>% 
  ungroup()
# It worked! No more duplicates

# Preparing RB data for merge
RB_data_c <- RB_data_c %>% 
  select(gameId, playId, DistFromBall_BS) %>% 
  rename(DistFromBall_RB = DistFromBall_BS)

# Merging with TE data
TE_data <- left_join(TE_data, RB_data_c, by = c("gameId", "playId"))
```

### TE Assignment

This will be the response variable for the model

It determines if a TE was running a route or blocking by:

1)  If the play is a designed rush, label: Block
2)  If the TE was running a route (wasRunningRoute == 1), label: Route
3)  If the TE didn't run a route and was recorded blocking a player, label: Block

```{r}
# How to determine if a TE was running a route or blocking?

# 1) if the play is a designed rush, label: Block
# 2) if the TE was running a route (wasRunningRoute == 1), label: Route
# 3) if the TE didn't run a route and was recorded blocking a player, label: Block

# Cleaning passResult so it will tell if the play was designed to be a pass or a rush
TE_data <- TE_data %>% 
  rename(playDesign = passResult) %>% 
  mutate(playDesign = case_when(
    playDesign == "" ~ "Rush",
    TRUE ~ "Pass"
  ))

# Converting significant NA values to 0
TE_data$blockedPlayerNFLId1[is.na(TE_data$blockedPlayerNFLId1)] <- 0
TE_data$wasRunningRoute[is.na(TE_data$wasRunningRoute)] <- 0

# Creating initial TE assignments
TE_data <- TE_data%>%
  mutate(TEassignment = case_when(
    playDesign == "Rush" ~ "Block",
    wasRunningRoute == 1 ~ "Route",
    qbSneak == TRUE ~ "Block",
    blockedPlayerNFLId1 > 0 ~ "Block",
    TRUE ~ "Other"
  ))

# Since there are only 730 observations that don't fit into either block or route, I'm going to drop them from the data set
TE_data <- TE_data %>% 
  filter(!(TEassignment == "Other"))
```

### Other Features Created

Total Yards to Go

```{r}
# Creating total yards to go variable
TE_data <- TE_data %>%
  mutate(totalYardsToGo = case_when(
    yardlineNumber == 50 ~ 50,
    possessionTeam == yardlineSide ~ (100 - yardlineNumber),
    TRUE ~ yardlineNumber
  ))
```

Red Zone

```{r}
# Creating if a team is in the redzone
TE_data <- TE_data %>%
  mutate(redZone = case_when(
    totalYardsToGo <= 20 ~ TRUE,
    TRUE ~ FALSE
  ))

# Removing qb spikes and kneels
TE_data <- TE_data %>% 
  mutate(qbSpike = case_when(
    is.na(qbSpike) ~ FALSE,
    TRUE ~ qbSpike
  ))

TE_data <- TE_data %>% 
  filter(!(qbKneel == 1)) %>% 
  filter(!(qbSpike == TRUE))
```

Score Differential

```{r}
# Creating variable for possessionTeamScore and oppTeamScore, and ScoreDiff
TE_data <- TE_data %>%
  mutate(
    possTeamScore = if_else(possessionTeam == homeTeamAbbr, preSnapHomeScore, preSnapVisitorScore),
    oppTeamScore = if_else(possessionTeam == homeTeamAbbr, preSnapVisitorScore, preSnapHomeScore)
  )

# Creating Variable for score differential
# If negative the poss team is behind, if positive the poss team is ahead
TE_data$scoreDiff <- TE_data$possTeamScore - TE_data$oppTeamScore
```

Converting gameClock

```{r}
# Converting time to readable interval
TE_data <- TE_data %>%
  mutate(gameTime = gameClock) %>% 
  separate(gameClock, into = c("minutes", "seconds"), sep = ":", convert = TRUE) %>% 
  mutate(seconds = round((seconds/60), 2)) %>% 
  mutate(gameClock = minutes + seconds)
```

If on LOS

```{r}
# Determining if the TE is on the LOS or not

TE_data <- TE_data %>% 
  mutate(OnLOS = case_when(
    DistFromLOS > 1.5 ~ FALSE, # If further than 1.5 yards from LOS
    DistFromLOS < LOS_Tackle ~ TRUE, # If they are closer to the LOS than their tackle
    (DistFromLOS - LOS_Tackle) <= 0.2 ~ TRUE, # If they are with 0.2 from their tackles LOS position
    DistFromLOS < 1 ~ TRUE, # Less than one yard from the LOS
    TRUE ~ FALSE
  ))
```

## Missingness

```{r}
# Dropping unnecessary variables
TE_data_c <- TE_data %>% 
  select(-c(yardlineSide, homeTeamAbbr, visitorTeamAbbr, yardlineNumber, minutes, seconds, preSnapHomeScore, preSnapVisitorScore, qbSneak, qbSpike, qbKneel, playAction, pff_runPassOption, hadRushAttempt, wasRunningRoute, wasTargettedReceiver, routeRan, blockedPlayerNFLId1, playDirection, Facing, super_key, possTeamScore, oppTeamScore))

# Visualizing missingness
TE_data_c %>% vis_miss()
```

Offensive formation and receiver alignment

```{r}
# For Offensive formation if it is labeled as Jumbo, Wildcat, or NA (formations with the least obs). I will label the formation as "OTHER"
# This will take care of the NA values

TE_data_c <- TE_data_c %>% 
  mutate(offenseFormation = case_when(
    is.na(offenseFormation) ~ "OTHER",
    offenseFormation == "JUMBO" |
    offenseFormation == "WILDCAT" ~ "OTHER",
    TRUE ~ offenseFormation
  ))

# For Receiver alignment if it is labeled as 3x0, 1x0, 1x1, 2x0, 4x1, NA (alignments with the least obs). I will label the alignment as "OTHER"
TE_data_c <- TE_data_c %>% 
  mutate(receiverAlignment = case_when(
    is.na(receiverAlignment) ~ "OTHER",
    receiverAlignment == "1x0" |
    receiverAlignment == "1x1" |
    receiverAlignment == "2x0" | 
    receiverAlignment == "3x0" |
    receiverAlignment == "4x1" ~ "OTHER",
    TRUE ~ receiverAlignment
  ))
```

pff_passCoverage and pff_manZone

```{r}
# Combine:
# Cover-1 Double -> Cover-1
# all Cover-3 variations -> Cover-3
# Cover-6 left and right -> Cover-6
# 2-man, Bracket, Miscellaneous, prevent, NA -> "Other"

TE_data_c <- TE_data_c %>% 
  mutate(pff_passCoverage = case_when(
    pff_passCoverage == "Cover-1 Double" ~ "Cover-1",
    pff_passCoverage == "Cover-3 Cloud Left" |
    pff_passCoverage == "Cover-3 Cloud Right" |
    pff_passCoverage == "Cover-3 Double Cloud" |
    pff_passCoverage == "Cover-3 Seam" ~ "Cover-3",
    pff_passCoverage == "Cover 6-Left" |
    pff_passCoverage == "Cover-6 Right" ~ "Cover-6",
    is.na(pff_passCoverage) |
    pff_passCoverage == "2-Man" |
    pff_passCoverage == "Bracket" |
    pff_passCoverage == "Miscellaneous" |
    pff_passCoverage == "Prevent" ~ "Other",
    TRUE ~ pff_passCoverage
  ))

# pff_manZone has almost zero effect on the outcome (TE assignment), so I'm removing it from the data set

ggplot(data = TE_data_c, aes(x = pff_manZone, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

TE_data_c <- TE_data_c %>% 
  select(-pff_manZone)
```

In motion

```{r, warning=FALSE}
ggplot(data = TE_data_c, aes(x = s_bs, fill = inMotionAtBallSnap)) +
  geom_density(alpha = 0.5) +
  xlim(0,1)

# Merging tracking_motion with TE data
TE_data_c <- merge(TE_data_c, tracking_motion, by = c("gameId", "playId", "nflId"))

# In motions criteria: max_s < 0.62 y/s, 0.4 seconds prior to ball snap, moved 1.2 yards (yard_chg > 1.2) with in 0.4 second window

TE_data_c <- TE_data_c %>% 
  mutate(inMotionAtBallSnap = case_when(
    is.na(inMotionAtBallSnap) & 
      max_s < 0.62 &
      yard_chg > 1.2 ~ TRUE,
    is.na(inMotionAtBallSnap) ~ FALSE,
    TRUE ~ inMotionAtBallSnap
  ))

ggplot(data = TE_data_c, aes(x = s_bs, fill = inMotionAtBallSnap)) +
  geom_density(alpha = 0.5) +
  xlim(0,1)

# The graphs shape hasn't change, so the code was implemented properly
```

Shifts

```{r, warning=FALSE}
# Criteria: the player moves more than 2.5 between lineset and ball snap
TE_data_c <- TE_data_c %>% 
  mutate(y_chg = abs(y_bs - y_ls))

ggplot(data = TE_data_c, aes(x = y_chg, fill = shiftSinceLineset)) +
  geom_density(alpha = 0.5) +
  xlim(0,1)

TE_data_c <- TE_data_c %>% 
  mutate(shiftSinceLineset = case_when(
    is.na(shiftSinceLineset) &
      (y_chg > 2.5) ~ TRUE,
    is.na(shiftSinceLineset) ~ FALSE,
    TRUE ~ shiftSinceLineset
  ))

ggplot(data = TE_data_c, aes(x = y_chg, fill = shiftSinceLineset)) +
  geom_density(alpha = 0.5) +
  xlim(0,1)

# The graphs shape hasn't change, so the code was implemented properly
```

Missing RB Distance

```{r}
# NA values are caused by no RB being present in the play, only 1.2% of TE data doesn't have a RB present
# Since the data isn't random missingness, bootstraping shouldn't be used
# I believe the best thing to do is give these RB Distance 0, this way we can keep the observations, while showing they aren't present
TE_data_c <- TE_data_c %>% 
  mutate(DistFromBall_RB = case_when(
    is.na(DistFromBall_RB) ~ 0.00,
    TRUE ~ DistFromBall_RB
  ))
```

Removing a few unimportant variables

```{r}
TE_data_c <- TE_data_c %>% 
  select(-c(x_bs, y_bs, s_bs, x_ls, y_ls, s_ls, x_fb, y_fb, s_fb))
```

No more missingness

```{r}
TE_data_c %>% vis_miss()
```

Saving resulting data

**Replace 'file' with your file directory**

```{r}
write.csv(TE_data_c, file = "file/TE data c.csv", row.names = FALSE)
```
