---
title: "Data Merging"
author: "Brendan McGuinness"
date: "2024-12-23"
output: html_document
---

# Loading and Merging Data Sets


```{r}
library(dplyr)
```

This notebook does the following:

  - Loads the raw data
  
  - Merges the data for feature engineering

*After downloading the raw data files from Kaggle, replace `file` with your data directory*
```{r}
games <- read.csv("file/games.csv")
player_play <- read.csv("file/player_play.csv")
players <- read.csv("file/players.csv")
plays <- read.csv("file/plays.csv")
tracking_week_1 <- read.csv("file/tracking_week_1.csv")
tracking_week_2 <- read.csv("file/tracking_week_2.csv")
tracking_week_3 <- read.csv("file/tracking_week_3.csv")
tracking_week_4 <- read.csv("file/tracking_week_4.csv")
tracking_week_5 <- read.csv("file/tracking_week_5.csv")
tracking_week_6 <- read.csv("file/tracking_week_6.csv")
tracking_week_7 <- read.csv("file/tracking_week_7.csv")
tracking_week_8 <- read.csv("file/tracking_week_8.csv")
tracking_week_9 <- read.csv("file/tracking_week_9.csv")
```

Getting player information for each relevant position
```{r}
players_RB <- players %>% 
  select(nflId, position, displayName) %>% 
  filter(position == "RB")

players_TE <- players %>% 
  select(nflId, position, displayName) %>% 
  filter(position == "TE")

players_T <- players %>% 
  select(nflId, position, displayName) %>% 
  filter(position == "T")

# Getting player_play information
player_play_c <- player_play %>% 
  select(gameId, playId, nflId, hadRushAttempt, inMotionAtBallSnap, shiftSinceLineset, wasRunningRoute, wasTargettedReceiver, routeRan, blockedPlayerNFLId1, blockedPlayerNFLId2, blockedPlayerNFLId3)

# Merging to each player position
RB_data <- merge(player_play_c, players_RB, by = "nflId")
TE_data <- merge(player_play_c, players_TE, by = "nflId")
T_data <- merge(player_play_c, players_T, by = "nflId")
```

Merging play and game data with TE data
```{r}
# Selecting appropriate columns
plays_c <- plays %>% 
  select(gameId, playId, quarter, down, yardsToGo, possessionTeam, yardlineSide, yardlineNumber, gameClock, preSnapHomeScore, preSnapVisitorScore, offenseFormation, receiverAlignment, passResult, qbSneak, qbKneel, qbSpike, playAction, pff_runPassOption, pff_passCoverage, pff_manZone)

# Getting game information
games_c <- games %>% 
  select(gameId, homeTeamAbbr, visitorTeamAbbr)

# plays and games merge
games_plays <- merge(games_c, plays_c, by = c("gameId"))

# Merging with TE data
TE_data <- merge(games_plays, TE_data, by = c("gameId", "playId"))
```

Tracking data
```{r}
# Combining all tracking data
tracking_all <- bind_rows(tracking_week_1, tracking_week_2, tracking_week_3, tracking_week_4, tracking_week_5, tracking_week_6, tracking_week_7, tracking_week_8, tracking_week_9)

# Player position at line set
tracking_players_ls <- subset(tracking_all, event == "line_set", select = c(gameId, playId, nflId, frameType, x, y, s))
# Player position at ball snap
tracking_players_bs <- subset(tracking_all, frameType == "SNAP", select = c(gameId, playId, nflId, frameType, x, y, s, playDirection))
# Ball position
tracking_ball <- subset(tracking_all, event == "line_set") %>% 
  filter(is.na(nflId))


# Removing nflId NAs (the football) observations from player tracking data
tracking_players_bs <- subset(tracking_players_bs, !is.na(nflId))
tracking_players_ls <- subset(tracking_players_ls, !is.na(nflId))


# Changing the names of the variables for players

# Before snap
tracking_players_bs <- tracking_players_bs %>%
  rename(
    x_bs = x,
    y_bs = y,
    s_bs = s
  )

# For the football (so it can be differentiated from the players)
tracking_ball <- tracking_ball %>%
  rename(
    x_fb = x,
    y_fb = y,
    s_fb = s
  )

# Line set
tracking_players_ls <- tracking_players_ls %>% 
  rename(
    x_ls = x,
    y_ls = y,
    s_ls = s,
  )

# Merging Tracking Data
# prep before merge
tracking_players_bs <- tracking_players_bs %>% 
  select(-frameType)

tracking_players_ls <- tracking_players_ls %>% 
  select(-frameType)

tracking_ball <- tracking_ball %>% 
  select(c(gameId, playId, x_fb, y_fb, s_fb))

# Merge
tracking_data <- merge(tracking_players_bs, tracking_players_ls, by = c("gameId", "playId", "nflId"))
tracking_data <- merge(tracking_data, tracking_ball, by = c("gameId", "playId"))

# Creating Distance from ball columns
# Distance from ball at line set
tracking_data <- tracking_data %>% 
  mutate(DistFromBall_LS = round(sqrt((abs(x_ls - x_fb))**2 + (abs(y_ls - y_fb))**2), digits = 2))

# Distance from ball at snap
tracking_data <- tracking_data %>%
  mutate(DistFromBall_BS = round(sqrt((abs(x_bs - x_fb))**2 + (abs(y_bs - y_fb))**2), digits = 2))

# Distance from line of scrimmage
tracking_data <- tracking_data %>% 
  mutate(DistFromLOS = round(abs(x_bs - x_fb), digits = 2))

# Distance from ball width (y-axis)
tracking_data <- tracking_data %>% 
  mutate(DistFromBall_Wid = round(abs(y_bs - y_fb), digits = 2))
```

Merging tracking data with positions
```{r}
TE_data <- merge(TE_data, tracking_data, by = c("gameId", "playId", "nflId"))
T_data <- merge(T_data, tracking_data, by = c("gameId", "playId", "nflId"))
RB_data <- merge(RB_data, tracking_data, by = c("gameId", "playId", "nflId"))
```

Saving resulting data
*Replace 'file' with your file directory*
```{r}
save(TE_data, file = "file/TE data.csv")
save(TE_data, file = "file/T data.csv")
save(TE_data, file = "file/RB data.csv")
```



