---
title: "EDA"
author: "Brendan McGuinness"
output: html_document
---

# EDA

This notebook will address

-   Analyzing significance of parameters
-   Exploring interactions

Load packages

```{r}
library(dplyr)
library(ggplot2)
library(gridExtra)
```

Load data

**Replace 'file' with your file directory**

```{r}
TE_data_c <- read.csv("file/TE data c.csv")
```

## Graphs

Down and yards to go

```{r}
first_d <- TE_data_c %>% 
  filter(down == 1) %>% 
  filter(yardsToGo != 10)

second_d <- TE_data_c %>% 
  filter(down == 2)

third_d <- TE_data_c %>% 
  filter(down == 3)

fourth_d <- TE_data_c %>% 
  filter(down == 4)

first_TE <- ggplot(data = first_d, aes(x = yardsToGo, fill = TEassignment)) +
  geom_density(alpha = 0.5)

second_TE <- ggplot(data = second_d, aes(x = yardsToGo, fill = TEassignment)) +
  geom_density(alpha = 0.5)

third_TE <- ggplot(data = third_d, aes(x = yardsToGo, fill = TEassignment)) +
  geom_density(alpha = 0.5)

fourth_TE <- ggplot(data = fourth_d, aes(x = yardsToGo, fill = TEassignment)) +
  geom_density(alpha = 0.5)

grid.arrange(first_TE, second_TE, third_TE, fourth_TE)

# For second down situations the line deciding whether a TE would mostly often block, or run a route, is around 7 yards to go
# On third and fourth down however, the line decreased to about 3 yards to go
```

Quarter

```{r}
ggplot(data = TE_data_c, aes(x = quarter, fill = factor(TEassignment))) +
  geom_bar(position = "fill") +
  labs(
    title = "TE Assignment per quarter",
    fill = "TE Assignment",
    y = "Factor",
    x = "Quarter"
  )

# Not much of a difference between the quarters, except in the 2nd quarter, TE tend to run more routes
```

Offensive Formation

```{r}
ggplot(data = TE_data_c, aes(x = offenseFormation, fill = TEassignment)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Offensive Formation Distribution",
    x = "Offensive Formation",
    y = "Count",
    fill = "TE Assignment"
  )

ggplot(data = TE_data_c, aes(x = offenseFormation, fill = factor(TEassignment))) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Factor Offensive Formation",
    x = "Offensive Formation",
    y = " Factor Percentage",
    fill = "TE Assignment"
  )

# Large dispersion between offensive formation and the TE assignment. When in an Empty or Shotgun formation, TEs more often block while another formation they mostly block
```

Receiver Alignment

```{r}
ggplot(data = TE_data_c, aes(x = receiverAlignment, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = receiverAlignment, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# Only 2x1 and 3x2 have an affect on the TE assignment
```

Pass Coverage

```{r}
ggplot(data = TE_data_c, aes(x = pff_passCoverage, fill = TEassignment)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = TE_data_c, aes(x = pff_passCoverage, fill = factor(TEassignment))) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# The pass coverage of the offense doesn't play a large role in determining the outcome. Except for goal line and 'other' category, goal line = block, other = route
```

In motion at ball snap

```{r}
ggplot(data = TE_data_c, aes(x = inMotionAtBallSnap, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = inMotionAtBallSnap, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# For the small amount TE that are in motion, it doesn't have a strong affect on TE assignment
```

Shift

```{r}
ggplot(data = TE_data_c, aes(x = shiftSinceLineset, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = shiftSinceLineset, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# Much greater significance that 'in motion' if there was a shift, the TE more often Blocks
```

Distance from Ball

```{r}
ggplot(data = TE_data_c, aes(x = DistFromBall_Wid, fill = TEassignment)) +
  geom_histogram()

ggplot(data = TE_data_c, aes(x = DistFromBall_Wid, fill = factor(TEassignment))) +
  geom_histogram(position = "fill") +
  xlim(1.3,25) +
  labs(
    title = "Factor Distance From Ball",
    x = "Yards away from ball",
    y = "Factor Percentage"
  )

# If they are less than 5 yards from the ball the TE most often Blocks, Further than 5 yards they run a route most of the time
```

Distance from ball density plots

```{r}
ggplot(data = TE_data_c, aes(x = DistFromBall_RB, fill = TEassignment)) +
  geom_density(alpha = 0.5) +
  xlim(1,15) +
  labs(title = "Significance of RB Distance",
       x = "Running Back Distance from Ball (Yards)",
       y = "Density",
       fill = "TE Role") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5))
```

Side of ball

```{r}
ggplot(data = TE_data_c, aes(x = SideBS, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = SideBS, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# No significance, will remove before modeling
```

Multiple TE and Number of TE

```{r}
# multiple TE
ggplot(data = TE_data_c, aes(x = multipleTE, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = multipleTE, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# If only one they are more often run a route, if there is more than one, then they more often block

# Number of TEs
ggplot(data = TE_data_c, aes(x = NumberofTE, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = NumberofTE, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# 2 and 3 have about the same influence on the TE's assignment. So for my model I'm only going to use 'multipleTE'
```

TE position

```{r}
ggplot(data = TE_data_c, aes(x = TEposition, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = TEposition, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# If not lined up as a TE, far more often runs a route
# The other position most often block
```

TE alignment

```{r}
ggplot(data = TE_data_c, aes(x = TEAlignment, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = TEAlignment, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# 0x0 most often Route, every other alignment is most often a block
```

Red zone

```{r}
ggplot(data = TE_data_c, aes(x = redZone, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = redZone, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# TEs block slightly more when in the red zone
```

On the LOS

```{r}
ggplot(data = TE_data_c, aes(x = OnLOS, fill = TEassignment)) +
  geom_bar()

ggplot(data = TE_data_c, aes(x = OnLOS, fill = factor(TEassignment))) +
  geom_bar(position = "fill")

# If on the LOS TE are slightly more often run a route, vise versa when not on the line
```

Score Differential

```{r}

ggplot(data = TE_data_c, aes(x = scoreDiff, fill = TEassignment)) +
  geom_histogram()

ggplot(data = TE_data_c, aes(x = scoreDiff, fill = factor(TEassignment))) +
  geom_histogram(position = "fill") +
  xlim(-30,30) +
  labs(
    title = "Factor Score Differential",
    x = "Score Differential",
    y = "Factor Percentage"
  )

# When behind TEs run routes more, but when ahead, they more often block
```

Distance from ball vs yards to go

```{r}
TE_1down <- TE_data_c %>% 
  filter(down == 1)

TE_2down <- TE_data_c %>% 
  filter(down == 2)

TE_3down <- TE_data_c %>% 
  filter(down == 3)

TE_4down <- TE_data_c %>% 
  filter(down == 4)

point_1 <- ggplot(data = TE_1down, aes(x = DistFromBall_Wid, y = yardsToGo, color = TEassignment)) +
  geom_point()

point_2 <- ggplot(data = TE_2down, aes(x = DistFromBall_Wid, y = yardsToGo, color = TEassignment)) +
  geom_point()

point_3 <- ggplot(data = TE_3down, aes(x = DistFromBall_Wid, y = yardsToGo, color = TEassignment)) +
  geom_point()

point_4 <- ggplot(data = TE_4down, aes(x = DistFromBall_Wid, y = yardsToGo, color = TEassignment)) +
  geom_point()

grid.arrange(point_1, point_2, point_3, point_4)
```

RB Distance & RB Distance vs Yards to Go

```{r}
# RB Distance
ggplot(data = TE_data_c, aes(x = DistFromBall_RB, fill = TEassignment)) +
  geom_histogram()

ggplot(data = TE_data_c, aes(x = DistFromBall_RB, fill = factor(TEassignment))) +
  geom_histogram(position = "fill") +
  labs(
    title = "Factor Score Differential",
    x = "RB Distance",
    y = "Factor Percentage"
  )

quar1 <- ggplot(data = TE_1down, aes(x = DistFromBall_RB, y = yardsToGo, color = TEassignment)) +
  geom_point()

quar2 <- ggplot(data = TE_2down, aes(x = DistFromBall_RB, y = yardsToGo, color = TEassignment)) +
  geom_point()

quar3 <- ggplot(data = TE_3down, aes(x = DistFromBall_RB, y = yardsToGo, color = TEassignment)) +
  geom_point()

quar4 <- ggplot(data = TE_4down, aes(x = DistFromBall_RB, y = yardsToGo, color = TEassignment)) +
  geom_point()

grid.arrange(quar1, quar2, quar3, quar4)

# RB Distance and yards to go
```

Distance from ball vs Score Differential

```{r}
# Seperating data by quarter, as well as first and 10 plays
TE_qtr_1 <- TE_data_c %>% 
  filter(quarter == 1)

TE_qtr_2 <- TE_data_c %>% 
  filter(quarter == 2)

TE_qtr_3 <- TE_data_c %>% 
  filter(quarter == 3)

TE_qtr_4 <- TE_data_c %>% 
  filter(quarter == 4)

qtr_1 <- ggplot(data = TE_qtr_1, aes(x = DistFromBall_Wid, y = scoreDiff, color = TEassignment)) +
  geom_point()

qtr_2 <- ggplot(data = TE_qtr_2, aes(x = DistFromBall_Wid, y = scoreDiff, color = TEassignment)) +
  geom_point()

qtr_3 <- ggplot(data = TE_qtr_3, aes(x = DistFromBall_Wid, y = scoreDiff, color = TEassignment)) +
  geom_point()

qtr_4 <- ggplot(data = TE_qtr_4, aes(x = DistFromBall_Wid, y = scoreDiff, color = TEassignment)) +
  geom_point()

grid.arrange(qtr_1, qtr_2, qtr_3, qtr_4)
```

## Exploring Interactions and Predictor Significance

```{r}
# Dropping variables that won't be used in the model or aren't worth analyzing
TE_data_cc <- TE_data_c %>% 
  select(-c(playDesign, position, DistFromBall_LS, DistFromBall_BS, DistFromLOS, TacklePos, LOS_Tackle, DistFromBall_Tackle, max_s, yard_chg, y_chg, gameTime))

# Changing all character variables into factors
TE_data_cc <- TE_data_cc %>% 
  mutate_if(is.character, as.factor)

# Changing downs and quarters into factors
TE_data_cc <- TE_data_cc %>% 
  mutate(
    quarter = as.factor(quarter),
    down = as.factor(down)
  )
```

```{r}
# Comparing all predictors and possible interactions
all_glm <- glm(TEassignment ~ . - gameId - playId - nflId - displayName - possessionTeam + down*yardsToGo + quarter*scoreDiff + DistFromBall_Wid*yardsToGo, data = TE_data_cc, family = binomial(link = "logit"))
anova(all_glm)
```

After analyzing all of these results, the features: `totalYardsToGo`, `redZone`, `NumberofTE`, `SideBS` won't be used for final modeling. They didn't show a great enough significance and therefore will create unnecessary noise in the model if left in.
