---
title: "Model Testing"
author: "Brendan McGuinness"
output: html_document
---

# Model Testing

This Notebook will address

-   Which models best fit to the overall TE data using ROC AUC values
-   Models include:
    -   Logistic Regression
    -   Elastic Net Regression
    -   Support Vector Machine
    -   Gradient Boosted Tree
    -   Random Forest

Load packages

```{r}
library(dplyr)
library(tidymodels)
library(ranger)
library(kableExtra)
library(glmnet)
library(ranger)
library(xgboost)
library(yardstick)
```

Load data

**Replace 'file' with your file directory**

```{r}
TE_data_c <- read.csv("file/TE data c.csv")
```

Feature dropping a data cleaning for model running

```{r}
TE_data_final <- TE_data_c %>% 
  select(-c(playDesign, position, x_bs, y_bs, s_bs, x_ls, y_ls, s_ls, x_fb, y_fb, s_fb, DistFromBall_LS, DistFromBall_BS, DistFromLOS, TacklePos, LOS_Tackle, DistFromBall_Tackle, max_s, yard_chg, y_chg, gameTime, possessionTeam, displayName))

# Changing all character variables into factors
TE_data_final <- TE_data_final %>% 
  mutate_if(is.character, as.factor)

# Removing features that don't show a strong enough significance to be used in the model
TE_data_final <- TE_data_final %>% 
  select(-c(totalYardsToGo, redZone, NumberofTE, SideBS))

# Changing downs and quarters into factors
TE_data_final <- TE_data_final %>% 
  mutate(
    quarter = as.factor(quarter),
    down = as.factor(down)
  )

# Changing all logistic variables into factors
TE_data_final <- TE_data_final %>% 
  mutate_if(is.logical, as.factor)
```

## Testing the Models

Splitting Data

```{r}
set.seed(1142)

# Splitting the data by 80% and stratifying by TEassignment
TE_split <- initial_split(TE_data_final, prop = 0.80, strata = TEassignment)

TE_train <- training(TE_split)
TE_test <- testing(TE_split)

# Stratified Sampling
TE_fold <- vfold_cv(TE_train, v = 5, strata = TEassignment)
```

Recipe which includes interaction required models

```{r}
TE_recipe <- recipe(TEassignment ~ ., data = TE_data_final) %>%
  step_interact(terms = ~ down:yardsToGo + quarter:scoreDiff + yardsToGo:DistFromBall_Wid) %>%
  step_center(all_numeric_predictors()) %>% 
  step_scale(all_numeric_predictors()) %>% 
  step_dummy(all_nominal_predictors())


prep(TE_recipe) %>%
  bake(new_data = TE_train) %>%
  head() %>%
  kable() %>% 
  kable_styling("striped", full_width = F) %>%
  column_spec(1:32, extra_css = "white-space: nowrap;") %>%
  row_spec(0, align = "c") %>%
  scroll_box(width = "100%")
```

Model setups

```{r}
# Logistic Regression
reg_set_TE <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

# Elastic Net Regression
net_set_TE <- logistic_reg(
  mixture = tune(),
  penalty = tune()) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification")

# Support Vector Machine (SVM)
svm_set_TE <- svm_rbf() %>% 
  set_engine("kernlab") %>% 
  set_mode("classification")

# Random Forest
rf_set_TE <- rand_forest(
  mtry = tune(),
  trees = tune(),
  min_n = tune()) %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")

# Gradient Boosted Tree
gbt_set_TE <- boost_tree(
  mtry = tune(),
  trees = tune(),
  learn_rate = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")
```

Work flows

```{r}
reg_wflow_TE <- workflow() %>% 
  add_model(reg_set_TE) %>% 
  add_recipe(TE_recipe)

net_wflow_TE <- workflow() %>% 
  add_model(net_set_TE) %>% 
  add_recipe(TE_recipe)

svm_wflow_TE <- workflow() %>% 
  add_model(svm_set_TE) %>% 
  add_recipe(TE_recipe)

rf_wflow_TE <- workflow() %>% 
  add_model(rf_set_TE) %>% 
  add_recipe(TE_recipe)

gbt_wflow_TE <- workflow() %>% 
  add_model(gbt_set_TE) %>% 
  add_recipe(TE_recipe)
```

Grids

```{r}
net_grid_TE <- grid_regular(penalty(),
                         mixture(range = c(0,1)),
                         levels = 10)

rf_grid_TE <- grid_regular(mtry(range = c(2,10)),
                             trees(range = c(200, 1000)),
                             min_n(range = c(2, 10)),
                             levels = 5)

gbt_grid_TE <- grid_regular(mtry(range = c(1,5)), 
                        trees(range = c(200, 1000)),
                        learn_rate(range = c(0.1, 0.3)),
                        levels = 5)
```

Tuning

```{r}
set.seed(1142)

net_tune_TE <- tune_grid(
  net_wflow_TE,
  resamples = TE_fold,
  grid = net_grid_TE
)

rf_tune_TE <- tune_grid(
  rf_wflow_TE,
  resamples = TE_fold,
  grid = rf_grid_TE
)

gbt_tune_TE <- tune_grid(
  gbt_wflow_TE,
  resamples = TE_fold,
  grid = gbt_grid_TE
)
```

Plots

```{r}
autoplot(net_tune_TE, metric = "roc_auc")
autoplot(rf_tune_TE, metric = "roc_auc")
autoplot(gbt_tune_TE, metric = "roc_auc")
```

Non-tuning models

```{r}
# Logistic Regression
reg_auc_TE <- TE_fold %>%
  mutate(
    fit = map(splits, ~ fit(reg_wflow_TE, data = analysis(.))),
    
    augmented = map2(fit, splits, ~ augment(.x, new_data = assessment(.y))),
    
    roc_auc = map(augmented, ~ roc_auc(.x, truth = TEassignment, .pred_Block))
  ) %>%
  select(roc_auc) %>%
  unnest(cols = c(roc_auc))

reg_auc_TE <- reg_auc_TE %>%
  group_by(.metric, .estimator) %>% 
  summarize(mean = mean(.estimate)) %>% 
  ungroup()

# SVM
svm_auc_TE <- TE_fold %>%
  mutate(
    fit = map(splits, ~ fit(svm_wflow_TE, data = analysis(.))),
    
    augmented = map2(fit, splits, ~ augment(.x, new_data = assessment(.y))),
    
    roc_auc = map(augmented, ~ roc_auc(.x, truth = TEassignment, .pred_Block))
  ) %>%
  select(roc_auc) %>%
  unnest(cols = c(roc_auc))

svm_auc_TE <- svm_auc_TE %>%
  group_by(.metric, .estimator) %>% 
  summarize(mean = mean(.estimate)) %>% 
  ungroup()
```

Best of tuned models

```{r}
best_net_TE <- show_best(net_tune_TE, metric = "roc_auc", n = 1)
best_rf_TE <- show_best(rf_tune_TE, metric = "roc_auc", n = 1)
best_gbt_TE <- show_best(gbt_tune_TE, metric = "roc_auc", n = 1)
```

Results between the 5 Models

```{r}
model_vector <- c('SVM', 'Logistic Regression', 'Random Forest', 'Gradient Boosted Tree', 'Elastic Net')
roc_results_TE <- bind_rows(svm_auc_TE, reg_auc_TE, best_rf_TE, best_gbt_TE, best_net_TE)
roc_results_TE$Models <- model_vector
roc_results_TE <- roc_results_TE[, c('Models', '.metric', 'mean')]
roc_results_TE[order(-roc_results_TE$mean), ]
```
